%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Deedy - One Page Two Column Resume
% LaTeX Template
% Version 1.1 (30/4/2014)
%
% Original author:
% Debarghya Das (http://debarghyadas.com)
%
% New author:
% Yubo Su (yssu@caltech.edu)
%
% Original repository:
% https://github.com/deedydas/Deedy-Resume
%
% IMPORTANT: THIS TEMPLATE NEEDS TO BE COMPILED WITH XeLaTeX
%
% This template uses several fonts not included with Windows/Linux by
% default. If you get compilation errors saying a font is missing, find the line
% on which the font is used and either change it to a font included with your
% operating system or comment the line out to use the default font.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[]{yubo-resume-openfont}

\begin{document}

\newcommand{\tab}{\hspace{10pt}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%     LAST UPDATED DATE
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\lastupdated

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%     TITLE NAME
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\namesection{Yubo}{Su}{\href{mailto:yssu@caltech.edu}{yssu@caltech.edu}
    $\bullet$ 770.527.2575\\
    \urlstyle{same}{http://www.linkedin.com/in/yubosu} $\bullet$
    \urlstyle{same}{https://github.com/yubo56}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%     COLUMN ONE
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{minipage}[t]{0.33\textwidth}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%     EDUCATION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Education}
    \subsection[Cornell]{Cornell University}
    \descript{Ph.D.~Astrophysics}
    \location{Jun 2016--Present \\
    Ithaca, NY}
    \sectionsep
    \subsection[CIT]{California Institute\\
        of Technology}
    \descript{B.S.~in Physics, \\
        \tab Computer Science}
    \location{Oct 2016--Jun 2016 \\
        Pasadena, CA | GPA:\@ 3.74}
\sectionsep

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%     SKILLS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Skills}
    \subsection{Programming}
    Javascript (Node.js) $\bullet$ Python $\bullet$ C/C++\\
    Java $\bullet$ Shell $\bullet$ CUDA $\bullet$ Assembly\\
    \sectionsep
    \subsection{Skillset}
    Numerical Simulation\\
    Systems Infrastructure \&\\
        \tab Optimization\\
    Data Management \& Security
    \sectionsep
    \subsection{Tools}
    Matlab $\bullet$ Mathematica $\bullet$ \LaTeX\\
    MongoDB $\bullet$ PostgreSQL \\
    AWS (EC2, S3, etc.) $\bullet$ Docker\\
    Ansible $\bullet$ Jenkins $\bullet$ Protractor\\
    Git $\bullet$ Linux\\
    \sectionsep
    \subsection{Languages}
    English $\bullet$ Chinese $\bullet$ French
    \sectionsep

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%     LINKS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% (ys) not using

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%     COURSEWORK
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Coursework}
    \subsection{Physics}
        Astrophysical Processes\\
        Advanced Plasma Physics\\
        Computational Physics\\
        Introduction to Particle Physics\\
        Introduction to Solid State Physics\\
    \sectionsep

    \subsection{Computer Science}
        Machine Learning\\
        GPU Programming\\
        Networks and Economics\\
        Relational Databases
    \sectionsep

    \subsection{Teaching}
        Differential Equations\\
        Complex Analysis\\
        C++ Language Workshop (1x)
    \sectionsep

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%     COLUMN TWO
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{minipage}
\hfill
\begin{minipage}[t]{0.66\textwidth}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%     EXPERIENCE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Experience}

\runsubsection{Blend Labs}
\descript{| Software Engineer}
\location{July 2016--Present | San Francisco, CA}
\vspace{\topsep} % Hacky fix for awkward extra vertical space
\begin{tightemize}
    \item Developed AWS S3 file management microservice. Implemented per-file
        encryption, set up load testing suite and stabilized all microservice
        deploys.
    \item Data permissioning to enable industry-first multiple accounts on
        single loan.
    \item Profiled and optimized unit tests and app deploy by parallelizing
        tests, improving build caching and decreasing app size. Average speed
        up of 2x.
    \item Stabilized unit and end-to-end tests, reducing failures by 3x to 99\%+
        stability.
    \item Improved scope/accessibility of data backup \& restore in app.
    \item Node.js, Angular, Mongo, Python, Docker, Shell, Ansible, AWS services
    \item Career fair recruiting, phone and on-site interviewing.
\end{tightemize}
\sectionsep

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%     RESEARCH
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Research}
\subsection{California Institute of Technology}
\descript{Undergraduate Researcher (Computational Cosmology)}
\location{Jan 2015--Jun 2016 | Pasadana, CA}
\begin{tightemize}
    \item Worked with \textbf{\href
            {http://www.astro.caltech.edu/~golwala/}
            {Prof Sunil Golwala}
        } to quantify detectability of kinetic Sunyaev-Zel'dovich Effect with
        future sub-millimeter telescopes.
    \item Developed a procedure to remove contaminating sources from
        simulated 2D telescope data conforming to accuracy bounds in optimal
        filtering theory.
    \item Code at
        \url{https://github.com/yubo56/Bolocam_Source_Subtraction}.
    \item Signal Processing, IDL, Linux
\end{tightemize}
\sectionsep

\subsection{NASA Jet Propulsion Laboratory}
\descript{Undergraduate Researcher (Solar Physics)}
\location{Jun 2014--Dec 2014 | Pasadena, CA}
\begin{tightemize}
    \item Worked with \textbf{\href
            {https://science.jpl.nasa.gov/people/Liewer/}
            {Dr.\ Paulett Liewer}
        } to generate synthetic white light images for solar phenomena
        simulating Solar Probe Plus (exp.~2020) view parameters.
    \item AGU 2014---
        {\small
            \url{https://agu.confex.com/agu/fm14/webprogram/Paper18882.html}
        }
    \item Raytracing, IDL, C
\end{tightemize}
\sectionsep

\subsection{California Institute of Technology}
\descript{Undergraduate Researcher (Computational Chemistry)}
\location{Mar 2013--Aug 2013 | Pasadena, CA}
\begin{tightemize}
    \item Worked with \textbf{\href
           {http://www.cce.caltech.edu/content/william-goddard}
           {Prof William A. Goddard}
        } and
        \textbf{\href
          {http://mendoza.eng.fsu.edu/}
          {Prof Jose L. Mendoza-Cortes}
        } to develop \emph{ab
        initio} force field parameters to compute hydrogen adsorption of
        various COF/MOFs.
    \item Numerical Optimization, Monte Carlo methods
\end{tightemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%     AWARDS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Awards}
\subsection{California Institute of Technology}
\begin{tabular}{p{20pt}p{80pt}p{7.5cm}}
    2016 & Best TA---Teaching Feedback & Among all Caltech Undergraduate and
    Graduate TAs. 22/24 students who responded gave perfect reviews in all
    categories.\\
    2016 & Outstanding Teaching Award & Nominated by students among teachers and
    TAs, selected by student body.\\
    2015 & NSF GRFP Honorable Mention & Proposed to study core-collapse
    supernovae gravitational waves using machine learning techinques.
\end{tabular}
\sectionsep

\end{minipage}
\end{document}
